I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-12-03 paper_io_sim.tex

			We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read and write mechanisms in Linux (Figure~\ref{fig:interaction}). 
			To read or write a file chunk, a simulated application sends a request 
			to the IOController. The IOController determines the amount of memory required by the operation,
			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
			and finally reads or writes on disk if necessary.
			The Memory Manager simulates flushing, eviction, and I/O to cache.

@@

			\paragraph*{Page cache LRU lists}
			In the Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages may 
			require significant computation overhead. 
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
   				\caption{Simulation of page cache LRU lists with data blocks}	\label{fig:lrulist}
			\end{figure}

			With the data block definition, page cache LRU lists can be modeled with 
			two lists of data blocks: an active list and an inactive list, ordered by 
			last access time (earliest first, see Figure~\ref{fig:lrulist}).
			As in the kernel, our simulator keeps the size of the active list under
			twice the size of the inactive list by moving least recently 
            used data blocks from the active list to the inactive list.
			Modeling page cache LRU lists as lists of data blocks reduces the
			overhead of the simulator, as data blocks are obviously less
			numerous than file pages, and at the same time achieves the accuracy in 
			I/O time simulation that can be seen in the later section.

@@

			\paragraph*{File read and write}
            Our model supports chunk-by-chunk file read and write simulation
            with a user-defined chunk size, in which chunks are read/written
            until file is entirely read/written.
            The simulated LRU lists are used as follows. 
            When a file is read, its data can be partially cached, completely cached, 
            or uncached. 
            In the scope of this study, our model support file accesses in chunk 
            but not file random access.
            Thus, we simplify the data read order, in which uncached data is read 
            from disk first, and cached data is only read after all uncached data is read 
            (left to right in Figure~\ref{fig:read_order}).
            When a chunk of uncached data is read, a new clean block is created 
            and put at the top of the inactive list. 
            For a chunk of cached data, one or more existing data blocks in two 
            page cache LRU lists are accessed again, and moved to the active list. 
			If the blocks are clean, we only need to keep the information about the 
			amount of data as they are available for eviction. 
			Thus, these re-accessed clean blocks can be merge into a bigger block 
			to reduce the number of block objects and put it in the active list. 
            If the blocks are dirty, since the blocks need to keep the information 
            of entry time for periodical flushing, they are only moved to the active list. 
			Because the chunk size and block size can be different, there are circumstances 
			in which a block is not entirely read. 
			In this case, the block is split into two smaller blocks, one is re-accessed 
			and the other remains the same.

>>>>

			We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read and write mechanisms in Linux (Fig.~\ref{fig:interaction}). 
			To read or write a file chunk, a simulated application sends a request 
			to the IOController. The IOController determines the amount of memory required by the operation,
			interacts with the Memory Manager accordingly, reads or write to cache through the Memory Manager, 
			and finally reads or writes on disk if necessary.
			The Memory Manager simulates flushing, eviction, and I/O to cache.

@@

			\paragraph*{Page cache LRU lists}
			In the Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages 
			induces substantial overheads.
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
   				\caption{Model of page cache LRU lists with data blocks.}	\label{fig:lrulist}
			\end{figure}

			We model page cache LRU lists as 
			two lists of data blocks, an active list and an inactive list, both ordered by 
			last access time (earliest first, Fig.~\ref{fig:lrulist}).
			As in the kernel, our simulator limits the size of the active list to
			twice the size of the inactive list, by moving least recently 
            used data blocks from the active list to the inactive list.
			Modeling page cache LRU lists as lists of data blocks reduces the
			overhead of the simulator, as data blocks are obviously less
			numerous than file pages, and at the same time achieves the accuracy in 
			I/O time simulation that can be seen in the later section.

@@

			\paragraph*{File read and write}
            Our model supports chunk-by-chunk file read and write simulation
            with a user-defined chunk size, in which chunks are read/written
            until file is entirely read/written.
            The simulated LRU lists are used as follows. 
            When a file is read, its data can be partially cached, completely cached, 
            or uncached. 
            In the scope of this study, our model support file accesses in chunk 
            but not file random access.
            Thus, we simplify the data read order, in which uncached data is read 
            from disk first, and cached data is only read after all uncached data is read 
            (left to right in Fig.~\ref{fig:read_order}).
            When a chunk of uncached data is read, a new clean block is created 
            and put at the top of the inactive list. 
            For a chunk of cached data, one or more existing data blocks in two 
            page cache LRU lists are accessed again, and moved to the active list. 
			If the blocks are clean, we only need to keep the information about the 
			amount of data as they are available for eviction. 
			Thus, these re-accessed clean blocks can be merge into a bigger block 
			to reduce the number of block objects and put it in the active list. 
            If the blocks are dirty, since the blocks need to keep the information 
            of entry time for periodical flushing, they are only moved to the active list. 
			Because the chunk size and block size can be different, there are circumstances 
			in which a block is not entirely read. 
			In this case, the block is split into two smaller blocks, one is re-accessed 
			and the other remains the same.
